---
title: "Practical Machine Learning Project"
author: "Roy Chen"
date: "Sunday, March 22, 2015"
output: html_document
---

##Summary of project:

The training data for this project are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: 
http://groupware.les.inf.puc-rio.br/har. 

In this project, I attempt to predict the manner in which people do exercise as supplied byt the training/testing data from above. 
I will describe how I built the model, used cross validation, and explain what I think the expected out of sample error is, and why I made the choices I did. 
I will also use the prediction model to predict 20 different test cases. 

##Data Processing:

```{r import files and processing, echo = TRUE}
library(lattice)
library(ggplot2)
library(caret)

training <- read.csv("pml-training.csv", header = TRUE, 
        stringsAsFactors = FALSE)

testing <- read.csv("pml-testing.csv", header = TRUE,
        stringsAsFactors = FALSE)

#str(training)
#training has 19622 obs, 160 variables

#set factors
training$user_name <- as.factor(training$user_name)
training$classe <- as.factor(training$classe)
training$cvtd_timestamp <- as.factor(training$cvtd_timestamp)
training$new_window <- as.factor(training$new_window)

#str(testing)
#testing has 20 obs, 160 variables

#set factors for testing
testing$user_name <- as.factor(testing$user_name)
testing$cvtd_timestamp <- as.factor(testing$cvtd_timestamp)
testing$new_window <- as.factor(testing$new_window)

#index of columns with NA's
cols.na <- apply(!is.na(training), 2, all)
training <- training[,cols.na] #down from 160 to 93 variables
testing <- testing[,cols.na] #93 as well

#removed column "X" which are the row numbers
training <- training[,-1]
testing <- testing[,-1]

#index of columns that contain numeric values
cols.ind <- sapply(training, is.numeric)

#preprocessing training:
pre.proc <- preProcess(training[,cols.ind], method = c("center", "scale", 
        "medianImpute"))
new.training <- predict(pre.proc, training[,cols.ind])
new.training <- cbind(user_name = training$user_name, cvtd_timestamp = 
        training$cvtd_timestamp, new_window = training$new_window,new.training)

#preprocessing testing set:
new.testing <- predict(pre.proc, testing[,cols.ind])
new.testing <- cbind(user_name = testing$user_name, cvtd_timestamp = 
        testing$cvtd_timestamp, new_window = testing$new_window,new.testing)
```

##Training Model:

Now that the training and testing sets have been processded accordingly, I proceed with building the model.

```{r model training, echo = TRUE}
set.seed(9876)

train.cont <- trainControl(method = "cv") #cross validation

fit <- train(training[,]$classe ~ ., method = "ctree", 
             data = new.training[,], trControl = train.cont)
```

Let's take a look at the fit results:

```{r model results, echo = TRUE}
fit$results
```

This shows that the out of sample error is 2.1% based on the best model with mincriterion = 0.50 and accuracy at 97.9%.

##Predicting of test set:

```{r predicting test sets, echo = TRUE}
predict.testing <- predict(fit, new.testing)
predict.testing
```


##Printing results for submission:

```{r submission, echo = TRUE}
#save answers as character vectors:
answers <- predict.testing

#from coursera project assignment page
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```